---
title: "Data Generator"
author: "Peng Chong"
date: "22/02/2019"
output: html_document
---
Heavily referenced from: https://xang1234.github.io/isochrone/
otp apis doc: http://dev.opentripplanner.org/apidoc/1.3.0/


Make sure otp server is up before running the code here
The codes here are for data generations

Setup:
```{r setup, include<-FALSE}
knitr::opts_chunk$set(echo <- TRUE)
packages = c('tidyverse','leaflet', 'knitr','stringr','httr','geojsonio', 'sp', 'dplyr', 'SpatialAcc', 'rgeos', 'spdplyr', 'KernSmooth', 'raster') 
for (p in packages){
  if(!require(p, character.only = T)){ 
    install.packages(p)
  }
  library(p,character.only = T) 
}

library(spdplyr)
#localhost ends WITHOUT /
localurl = 'http://localhost:8080'

#Respective data paths ends WITHOUT /
isopath = 'data/isodata'
csvpath = 'data/csvdata'
matrixpath = 'data/matrixdata'
spatialpath = 'data/spatialdata'
objpath = 'data/objdata'

all.postals = read_csv(paste0(csvpath, '/all_postal_code.csv'))
all.residential = read_csv(paste0(csvpath, '/RESIDENTIAL-prepared-hdb-details.csv'))
schs.infos = read_csv(paste0(csvpath, '/general-information-of-schools.csv'))
schs.infos$postal_code =  as.numeric(schs.infos$postal_code)

schs.infos = inner_join(schs.infos, all.postals, by = c('postal_code' = 'POSTAL')) %>% dplyr::select('school_name', 'postal_code', 'LATITUDE', 'LONGITUDE', 'X', 'Y', 'ROAD_NAME') %>% group_by_at(vars( postal_code)) %>% filter(row_number() == 1)


residential.infos = inner_join(all.residential, all.postals, by = c('POSTAL' = 'POSTAL')) %>% dplyr::select('ADDRESS' = 'ADDRESS.x', 'POSTAL', 'LATITUDE' = 'LATITUDE.x', 'LONGITUDE' = 'LONGITUDE.x', 'X' = 'X.x', 'Y' = 'Y.x', 'ROAD_NAME' = 'ROAD_NAME.x') %>% group_by_at(vars( POSTAL)) %>% filter(row_number() == 1)


#fixing errors 
residential.infos = residential.infos[-c(3724),]

crsobj = CRS("+init=EPSG:3857")
coordinates(schs.infos)<-~LONGITUDE+LATITUDE
proj4string(schs.infos) = crsobj

coordinates(residential.infos)<-~LONGITUDE+LATITUDE
proj4string(residential.infos) = crsobj



```

Save to file
```{r}
saveRDS(all.postals, file='data/objdata/all.postal.rds')
saveRDS(residential.infos, file='data/objdata/all.residential.rds')
saveRDS(schs.infos, file='data/objdata/all.school.rds')
```



functions
```{r}

#Getting the data of a trip
get_dist = function(fromLat, fromLong, toLat, toLong){
    current <- GET(
      paste0(localurl, '/otp/routers/default/plan'),
      query = list(
        fromPlace = paste(fromLat,fromLong,sep = ","),
        toPlace = paste(toLat,toLong,sep = ","),
        mode = "WALK,TRANSIT",
        date = "03-27-2019",
        time= "07:00am",
        maxWalkDistance = '1600'
      )
    )
    #return(content(current))
    if (length(content(current)[["plan"]][["itineraries"]][[1]]) == 0){
      return(NULL)
    }else{
      return(content(current)[["plan"]][["itineraries"]][[1]])
    }
    
}

#Generating the dist and duration from all residential HDB and storing it for each school
generate_dist <-function(residential.info, destLat, destLong, schName){
  result = data.frame(matrix(ncol = 6, nrow= 0))
  colnames(result) = c('address', 'destination', 'latitude', 'longitude', 'duration', 'distance')
  
  for(i in 1: length(residential.info@coords[,1])){
    lat = as.double(residential.info@coords[i,]['LATITUDE'])
    long = as.double(residential.info@coords[i,]['LONGITUDE'])
    address = residential.info@data[i,]$ADDRESS
      info = get_dist(lat, long, destLat, destLong)
      if (is.null(info)){
        next()
      }
      total = 0
      for (j in 1:length(info$legs)){
        total = total + info$legs[[j]]$distance
      }
      
      dist = total/1000
      dur = info$duration/60
      result[nrow(result)+1,] = list(address, schName, lat, long, dur, dist)
      
  }
  saveRDS(result, file = paste0(matrixpath, '/', schName, '.rds'))
  return(result)
}

#isochrone generation and storing it for each school
get_geojson<-function(lat,lng,filename, rb = FALSE){
    current <- GET(
       paste0(localurl, '/otp/routers/current/isochrone'),
      query = list(
        fromPlace = paste(lat,lng,sep = ","),
        mode = "WALK,TRANSIT",
        date = "03-27-2019",
        time= "07:00am",
        geoidElevation = TRUE,
        maxWalkDistance = 1600,
        walkReluctance = 5,
        minTransferTime = 60,
        cutoffSec = 900,
        cutoffSec = 1800,
        cutoffSec = 2700,
        cutoffSec = 3600,
        cutoffSec = 5400
      )
    )
    current <- content(current, as = "text", encoding = "UTF-8")
    write(current, file = paste0(isopath, '/', filename,".geojson"))
    
    if(rb){
      sp = geojsonio::geojson_read(paste0(isopath, '/', filename,".geojson"), what = "sp")
      file.remove(paste0(isopath, '/', filename,".geojson"))
      return(sp)
    }
}

#used for internal testing and checking of isochrone
show_isochrone = function(sch,path, lng, lat, res_sp){
iso <- geojsonio::geojson_read(path,
  what = "sp")

pal=c('bisque','cyan','gold','tomato','red')

m =leaflet() %>%
  setView( lng = 103.8198, lat = 1.3521, zoom = 11) %>%
  addProviderTiles(providers$CartoDB.DarkMatter,
                   options = providerTileOptions(opacity = 0.8)) %>%  
  addPolygons(data =iso, stroke = TRUE, weight=0.5,
              smoothFactor = 0.3, color="black",
              fillOpacity = 0.2,fillColor =pal ) %>%
  addCircleMarkers( lng = residential.infos@coords[,1], lat = residential.infos@coords[,2], opacity = 1, fillOpacity = 1, color = '#ADFF2F', stroke=FALSE, radius= 2, popup = residential.infos@data$ADDRESS, label = sch, data = sch) %>%
  addMarkers(lng = lng, lat = lat, popup = sch, options = markerOptions(interactive = TRUE), clusterOptions = markerClusterOptions()) %>%
  addLegend(position="bottomleft",colors=rev(c("cornsilk", "lightskyblue","greenyellow","gold","tomato")),
            labels=rev(c("90 min", "60 min","45 min",
                     "30 min","15 min")),
            opacity = 0.8,
            title=paste0("Travel Time with Public Transport to ", sch) ) %>%
  addLegend(position="bottomright",colors=rev(c('greenyellow')),
            labels=rev(c("HDB")),
            opacity = 0.8) %>%
  htmlwidgets::onRender("
            function(el,x) {
                map = this;
            }
        ")
return(m)
}

#applying hansen to the stored data generated
apply_hansen = function(res){
  result = res %>% mutate(demand =100)
  durationMatrix = result %>% dplyr::select('duration')
  acc_hansen <- data.frame(ac(result$demand,
               c(10), durationMatrix,
               power = 0.01, family = "Hansen"))
  result_duration = result
  result_duration$durationHansen = acc_hansen[,1]
  result_duration$durationHansen[is.na(result_duration$durationHansen)] = 0
  
  
  distanceMatrix = result %>% dplyr::select('distance')
  acc_hansen <- data.frame(ac(result$demand,
               c(10), distanceMatrix,
               power = 0.01, family = "Hansen"))
  result_final = result_duration
  result_final$distanceHansen = acc_hansen[,1]
  result_final$distanceHansen[is.na(result_final$distanceHansen)] = 0
  # colnames(result_final)[8:9] = c('durationHansen', 'distanceHansen')
  return (result_final)
}

#hansen has to be named as accHansen on the spdf
#type is either duration or distance
#applying hansen calculations into the contours generated
integrate_hansen = function(contour.poly, sp.points.poly, type){
  uni = rev(sort(unique(contour.poly$Value)))
  points.copied = sp.points.poly
  contour.copied = contour.poly
  if (type == 'duration'){
      contour.copied$duration_hansen = 0
      
      for(j in 1:length(uni)){
        polys = contour.copied[contour.copied$Value == uni[j],]
        for(i in 1:nrow(polys)){
          ff = gContains(polys[i,], points.copied, byid = TRUE)
          points.copied$checker = ff
          sub = points.copied %>% filter(checker == TRUE)
          total = sum(sub$durationHansen, na.rm = TRUE)
          avg = total/length(sub)
            if (is.na(avg)){
              avg = 0
            }
          polys@data$duration_hansen[i] = avg
          points.copied = points.copied %>% filter(checker == FALSE)
        }
        temp.df = contour.copied %>% filter(Value == uni[j]) %>% mutate(duration_hansen = polys$duration_hansen) 
        temp.df = temp.df@data
      
        for(i in 1:nrow(temp.df)){
          f = which(contour.copied@data$area_sqkm == temp.df$area_sqkm[i], arr.ind=TRUE)
          contour.copied@data$duration_hansen[f] = temp.df$duration_hansen[i]
        }
      }
      return(contour.copied)
  }else{
      contour.copied$distance_hansen = 0
      
      for(j in 1:length(uni)){
        polys = contour.copied[contour.copied$Value == uni[j],]
        for(i in 1:nrow(polys)){
          ff = gContains(polys[i,], points.copied, byid = TRUE)
          
          points.copied$checker = ff
          sub = points.copied %>% filter(checker == TRUE)
          total = sum(sub$distanceHansen, na.rm = TRUE)
          avg = total/length(sub)
            if (is.na(avg)){
              avg = 0
            }
          polys@data$distance_hansen[i] = avg
          points.copied = points.copied %>% filter(checker == FALSE)
        }
        temp.df = contour.copied %>% filter(Value == uni[j]) %>% mutate(distance_hansen = polys$distance_hansen) 
        temp.df = temp.df@data
      
        for(i in 1:nrow(temp.df)){
          f = which(contour.copied@data$area_sqkm == temp.df$area_sqkm[i], arr.ind=TRUE)
          contour.copied@data$distance_hansen[f] = temp.df$distance_hansen[i]
        }
      }
      return(contour.copied)
  }
  
}

```

```{r}
#internal - correcting colnames
for (i in 1:2){
  lsch = schs.infos@data[i,'school_name']
  res = readRDS(paste0('data/matrixdata/',lsch,'.rds'))
  colnames(res) = c('address', 'destination', 'latitude', 'longitude', 'duration', 'distance')
  saveRDS(res, file = paste0('data/matrixdata/',lsch,'.rds'))

}
```



```{r}
#internal - converting to sp
for(i in 1: 357){
  print(i)
  lsch = schs.infos@data[i,'school_name']
  res = readRDS(paste0('data/matrixdata/',lsch,'.rds'))
  coordinates(res) = ~ longitude +  latitude
  proj4string(res) = proj4string(residential.infos)
  saveRDS(res, file = paste0('data/matrixdata/',lsch,'.rds'))
}
```


Generation of contouring of kernel density of residential HDB and storing it
```{r}
d2d = bkde2D(cbind(residential.infos@coords[,1],residential.infos@coords[,2]),bandwidth=c(0.0025,0.0025))
contour(d2d$x1,d2d$x2,d2d$fhat)
lines = contourLines(x=d2d$x1,y=d2d$x2,z = d2d$fhat,nlevels = 8)
dd1 = sapply(1:length(lines),function(i) Polygon(as.matrix(cbind(lines[[i]]$x,lines[[i]]$y))))
dd2 = sapply(1:length(lines),function(i) Polygons(list(dd1[[i]]),i))
poly_data = data.frame(Value = sapply(1:length(lines),function(i) lines[[i]]$level))
dd3 = SpatialPolygonsDataFrame(SpatialPolygons(dd2),data = poly_data)
values = unique(sapply(1:length(dd3@data[["Value"]]),function(i) dd3@data[["Value"]]))[,1]
dd3$area_sqkm <- area(dd3) / 1
proj4string(dd3) = proj4string(residential.infos)
saveRDS(dd3, file='data/objdata/kernel.residential.rds')
```

Testing out contour overlay
```{r}
pal <- colorFactor(
          palette = 'Greens',
          domain = result_acc$accHansen
        )

dd3 = readRDS(file='data/objdata/kernel.residential.rds')

uni = unique(dd3$Value)

m = leaflet() %>% addProviderTiles('CartoDB.DarkMatter', group = 'providertiles')
for(j in 1:length(uni)){
  polys = dd3[dd3$Value == uni[j],]
  m = m %>% addPolygons(data =polys, stroke = TRUE, weight=0.5, smoothFactor = 0.2, color="black", fillOpacity = 0.6, fillColor = brewer.pal(length(uni),"Greens")[j] )
}
m
```


Here generates the isochrone maps and the distance/duration data for all the residential HDB to each school

```{r}
#isochrone generation for all school - WARNING IO intensive, may take to an hour to run
for (i in 1:357){
  lsch = schs.infos@data[i,'school_name']
  print(lsch)
  lat = schs.infos@coords[i,2]
  long = schs.infos@coords[i,1]
  print(lat)
  print(long)

 iso = get_geojson(lat, long, schs.infos@data[i,'school_name'], rb = TRUE)
 proj4string(iso) = proj4string(residential.infos)
 res = colSums(gContains(iso, residential.infos, byid = TRUE))
 res = tbl_df(res)
 colnames(res) = 'blocks'
 iso@data = bind_cols(iso@data, res)
 saveRDS(iso, file = paste0(isopath, '/', lsch, '.rds'))
```

```{r}
#hansen generation for all school - WARNING IO intensive, may take up to a week to run
for (i in 1:357){
  lsch = schs.infos@data[i,'school_name']
  res = readRDS(paste0(matrixpath, '/',lsch,'.rds'))
  result_acc = apply_hansen(res)
  saveRDS(result_acc, paste0(matrixpath, '/',lsch,'.rds'))
  coordinates(result_acc) = ~ longitude +  latitude
  proj4string(result_acc) = proj4string(residential.infos)
  
  semi = integrate_hansen(dd3, result_acc, 'duration')
  full = integrate_hansen(semi, result_acc, 'distance')
  saveRDS(full, paste0(matrixpath, '/',lsch,'.contour.rds'))
  
}
```

